---
output: md_document
---

[![Build Status](https://travis-ci.org/ukgovdatascience/eesectors.svg?branch=master)](https://travis-ci.org/ukgovdatascience/eesectors)
[![Build status](https://ci.appveyor.com/api/projects/status/vulmerft4p30339l/branch/master?svg=true)](https://ci.appveyor.com/project/ivyleavedtoadflax/eesectors/branch/master)
[![codecov.io](http://codecov.io/github/ukgovdatascience/eesectors/coverage.svg?branch=master)](http://codecov.io/github/ukgovdatascience/eesectors?branch=master)
[![GitHub tag](https://img.shields.io/github/tag/ukgovdatascience/eesectors.svg)](https://github.com/ukgovdatascience/eesectors/releases)

```{r setup, echo=FALSE, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE
  )

```

# eesectors
## Functions for producing the Economic Estimates for DCMS Sectors First Statistical Release

**This is a prototype and subject to constant development**

This package provides functions used in the creation of a Reproducible Analytical Pipeline (RAP) for the Economic Estimates for DCMS sectors publication.

See the [RAP-demo-md](https://github.com/ukgovdatascience/RAP-demo-md) repository for an example of implementing these functions in the context of a Statistical First Release (SFR).

## Installation

The package can then be installed using `devtools::install_github('ukgovdatascience/eesectors')`. 
Some users may not be able to use the `devtools::install_github()` commands as a result of network security settings. 
If this is the case, `eesectors` can be installed by downloading the [zip of the repository](https://github.com/ukgovdatascience/govstyle/archive/master.zip) and installing the package locally using devtools::install_local(<path to zip file>).

## Quick start

This package provides functions to recreate Chapter three of the [Economic estimates of DCMS Sectors](https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/544103/DCMS_Sectors_Economic_Estimates_-_August_2016.pdf).

The package is built around the `long_data` class.
To create a `long_data` object, a `data.frame` must be passed to it which contains all the data required to produce the tables and charts in Chapter three.

An example of how this dataset will need to look is bundled with the package: `GVA_by_sector_2016`.
These data were extracted directly from the 2016 SFR which is in the public domain, and provide a test case for evaluating the data.

```{r}
library(eesectors)
GVA_by_sector_2016
```

When an object is instantiated into the `long_data` class, a number of checks are run on the data passed as the first argument.
These are explained in more detail in the help `?long_data()`

```{r warning=TRUE, message=TRUE}
gva <- long_data(GVA_by_sector_2016)
```

Any failed checks are raised as warnings, not errors, and so the user is able to continue.
However it is also possible to log these warnings as github issues by setting `log_issues=TRUE`.
This is a prototype feature that needs additional work to increase the usefulness of these issues, see below for details on environmental variables that are required for this functionality to work.

### Creating tables and charts

Tables and charts for Chapter three can be reproduced simply by running the relevant functions:

```{r}
format_table(gva)
```

```{r figure3.1}
figure3.1(gva)
```

Note that figures produced remain `ggplot2` objects, and can therefore be edited in the following way:

```{r figure3.2}
p <- figure3.2(gva)

p
```



```{r figure3.2edited}
library(ggplot2)

p + 
  ggtitle('Figure 3.2')
```

Note that figures make use of the [govstyle](https://github.com/ukgovdatascience/govstyle) package.
See the [vignette](https://github.com/ukgovdatascience/govstyle/blob/master/vignettes/absence_statistics.md) for more information on how to use this package.

### Logging data issues to github

In order to use this functionality, it is necessary to set the three following environmental variables:

|Name|Example|Description|
|---|---|---|
|GITHUB_PAT|_|A github personal access token with the necessary permissions.|
|LOG_REPO|RAP-demo-md|The name of a github repository where data issues can be logged.|
|LOG_OWNER|ukgovdatascience|The owner of the repository referred to in LOG_REPO.|

Environmental variables can be set interactively using `Sys.setenv()`, or more permanently by settin gteh variables in an `.Renviron` file which will be sourced when the project is loading (assuming you are using projects within Rstudio).
