---
title: "Calculate aggregated GVA figures"
output:
  html_notebook: default
  html_document: default
---

This notebook does two things:

* Extracts the data from the [../../OFFICIAL/OFFICIAL_working_file_dcms_V13.xlsm](../../OFFICIAL/OFFICIAL_working_file_dcms_V13.xlsm) and saves them out into `*.Rds` files.
* Combines data from the several data sources into a single dataset.
* Tests this single dataset for consistency with the existing 2016 figues, and makes multiple tests throughout the process.

### Setup

```{r setup, echo=FALSE}

devtools::install_github('ukgovdatascience/eesectors')
library(eesectors)
#devtools::load_all()


knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  error = FALSE
)

```

```{r load_packages}

# Note that in places I used :: to avoid confusion: there are duplicated
# functions with common sames like select and filter

library(dplyr)
library(tidyr)
library(testthat)
#library(eesectors)

```

First, load the data extracted from the spreadsheet `OFFICIAL_working_file_dcms_V13.xlsm`. These are stored as `*.Rds` files within the `../../OFFICIAL/cleaned/` folder.
Note that you must remove the password from this file for this to work.

```{r load_data}

# These data come from the worksheets `1997 Use`:`2015 Use`.

input <- '~/Documents/DSprojects/eesectors_working/working_file_dcms_V13.xlsm'

ABS <- extract_ABS_data(input)
# DCMS_sectors <- extract_DCMS_sectors(input): for now use data built into package
GVA <- extract_GVA_data(input)
SIC91 <- extract_SIC91_data(input)
tourism <- extract_tourism_data(input)
 
```

# Recreating the Working File worksheet 

To recreate the working file as it appears in the spreadsheet we must:

1. Deal with the SIC 91 data (i.e. remove SIC 91 data from `ABS` and swap in values from `SIC91)
2. Duplicate the 2014 ABS values to use for 2015 (2015 values not being available)
3. Merge the DCMS `sectors` to `ABS_2015` to get the 2 digit SIC.
4. Calculate sums across sectors and years.
5. Add in total UK GVA from `gva`.
6. Match in Tourism data.
7. Add tourism overlap.
8. Build the `GVA_by_sector` dataframe expected by the `eesectors` package.
9. Test the dataframe against the `eesectors::GVA_by_sector_2015`.
10. Print differences between `GVA_by_sector` and `eesectors::GVA_by_sector_2015`.

## 1. Deal with SIC 91

The `SIC91` data needs to be merged into `ABS` as the first operation. 
This replaces the SIC 91 data from the Annual Business Survey (ABS) with sales data.

```{r merge_SIC91_and_ABS}

# Remove all the rows that correspond to SIC 91 based on the unique SIC codes in
# SIC91.

SIC91_mask <- unique(SIC91$SIC)

SIC91_mask

ABS_no91 <- ABS %>% dplyr::filter(!SIC %in% SIC91_mask)

ABS_no91 

```
Our expectation is now that there are not any 91 SIC codes in the ABS dataset that are also classed as DCMS sectors. 

Test this here:

```{r test_merge_SIC91_and_ABS}
test_that(
  'sectors with 91 SIC 2 will match properly.',
  {
    
    # Create objects that contain just the rows which have SIC codes beginning
    # with 91 (there are some remaining even after the above step)
    
    sectors_91 <- DCMS_sectors$SIC[grepl('^91', DCMS_sectors$SIC)]
    ABS_91 <- ABS_no91$SIC[grepl('^91', ABS_no91$SIC)]
    
    # Check that none of the remaining 91 values in ABS are going to get matched
    # against DCMS sectors in sectors:
    
    expect_equal(0, sum(sectors_91 %in% ABS_91))
    
    # Check that when we merge ABS with SIC91, the new SIC91 entries are going 
    # to match the SIC codes for DMCS sectors that are in sectors. Note that not
    # all of the SIC4 codes in SIC91 are present in DCMS sectors, so check the 
    # other way, i.e. all SIC codes starting 91 in sectors$SIC are present in 
    # SIC91$SIC
    
    expect_equal(length(sectors_91), sum(sectors_91 %in% SIC91$SIC))
    
    }
  )

```

Now merge `ABS_no91` with `SIC91`.

```{r merge_ABS_without_91_and_SIC91}
ABS_91 <- dplyr::full_join(
  ABS_no91,
  SIC91
)

ABS_91

```

Test that this merge worked.

```{r test_merge_ABS_without_91_and_SIC91}

test_that(
  'merge of ABS_91 and SIC91 worked as expected.',
  {
    
    # Our expectation is that there should now be the same number of rows in
    # ABS91 as ABS, since we have replaced the 91 codes.
    
    expect_equal(nrow(SIC91), nrow(ABS) - nrow(ABS_no91))
    expect_equal(nrow(ABS), nrow(ABS_91))
    
    # Also we expect that there are valid abs values for these rows!
    # Extract out those rows:
    
    merged_rows <- ABS_91$ABS[grepl('^91\\.',ABS_91$SIC)]
    
    expect_true(all(!is.na(merged_rows)))
    expect_true(all(merged_rows > 0))
    
    }
  )


```

### 2. Duplicate 2014 ABS data for 2015

Since the `ABS_91` data only run to 2014, first duplicate the 2014 to be used for 2015:

```{r duplicate_2014_ABS}
# duplicate the 2014 data for 2015

ABS_2015 <- ABS_91 %>%
  dplyr::filter(year == 2014) %>%
  dplyr::mutate(year = 2015) %>%
  dplyr::bind_rows(ABS_91)

ABS_2015 

```

```{r test_duplicate_2014_ABS}
test_that(
  '2014 ABS data was replicated as 2015.',
  {
    
    # Check that the 2014 and 2015 ABS GVA are identical for 2014 and 2015
    
    GVA_2014 <- ABS_2015 %>%
      dplyr::filter(year == 2014)
    
    GVA_2015 <- ABS_2015 %>%
      dplyr::filter(year == 2015)
    
    expect_equal(GVA_2014$ABS_ind_GVA, GVA_2015$ABS_ind_GVA)
    
  }
)

```

### 3. Merge sectors and ABS datasets

```{r merge_sectors_and_ABS_2015}

GVA_sectors <- DCMS_sectors %>%
  dplyr::left_join(ABS_2015) %>%
  dplyr::rename(ABS_ind_GVA = ABS)

GVA_sectors

```

```{r test_merge_sectors_and_ABS_2015}

test_that(
  'merge of sectors and ABS_2015 worked as expected',
  {
    
    # Check that all of the columns from sectors are now in GVA
    
    expect_true(all(colnames(DCMS_sectors) %in% colnames(GVA_sectors)))
    
    # Check that all of the columns from ABS_2015 are now in GVA. Note
    # that we renamed abs, so we should check for that too.
    
    expect_equal(sum(colnames(ABS_2015) %in% colnames(GVA_sectors)), 3)
    
    # Check that there aren't a crazy amount of NAs in the dataframe
    
    expect_equal(c('DOMVAL', 'ABS', 'SIC'), na_cols(ABS_2015))
    
    # Check the proportion of these cols that are NA. NOte that these values are
    # somewhat arbitrary. In future years these will need to be reset.
    
    expect_lt(sum(is.na(ABS_2015$SIC))/nrow(ABS_2015), 0.01)
    
    expect_lt(sum(is.na(ABS_2015$DOMVAL))/nrow(ABS_2015), 0.01)
    
    expect_lt(sum(is.na(ABS_2015$abs))/nrow(ABS_2015), 0.2)
    
  }
)

```

Then calculate the 2 digit SIC total GVA (from `ABS_91`) for each of the DCMS sectors.
Extract unique 2 digit SICs:

```{r get_SIC2_unique}

# Extract all the unique 2 digit SICs

SIC2_unique <- GVA_sectors %>%
  dplyr::filter(!is.na(SIC2)) %$%
  SIC2 %>%
  unique

SIC2_unique

```

We expect there to be 23 of these, so run a quick check:

```{r test_get_SIC2_unique}

test_that(
  'there are the right number of unique SIC2s',
  {
    expect_equal(length(SIC2_unique), 23)
  }
)
```

Next we use `SIC2_unique` to extract the 2 digit SIC totals from `ABS_91`.

```{r calculate_SIC_2_totals_from_ABS}

# This will form the denominator in our division

denom <- ABS_2015 %>%
  dplyr::filter(SIC %in% SIC2_unique) %>%
  dplyr::mutate(ABS_2digit_GVA = ABS) %>%
  dplyr::select(year, SIC2 = SIC, ABS_2digit_GVA)

denom

```

Now join this back into `GVA`.

```{r merge_SIC_2_totals_into_GVA}

# Join back into GVA for division

GVA_sectors <- GVA_sectors %>% 
  dplyr::filter(present) %>%
  dplyr::left_join(denom) %>%
  dplyr::mutate(
    perc_split = ABS_ind_GVA / ABS_2digit_GVA
  )

# At this point there will be two rows that contain NAs, both as a result of
# tourism, so check for and drop these.

GVA_sectors <- GVA_sectors %>% 
  dplyr::filter(!(is.na(year) & is.na(ABS_ind_GVA)))

GVA_sectors

```

Test that this worked correctly.

```{r test_merge_GVA_with_denom}

test_that(
  'there are no unexpectedly missing values in GVA.',
  {
    
    # See http://stackoverflow.com/questions/10574061/show-columns-with-nas-in-a-data-frame
    
    na_cols <- function(df) colnames(df)[unlist(lapply(df, function(x) any(is.na(x))))]
    
    # We only expect NAs in DOMVAL at this point
    
    expect_equal('DOMVAL', na_cols(GVA_sectors))
  }
)

```

Now, multiply `perc_split` by `GVA` after joining with `GVA` to get the `BB16_GVA` (column Q in `Working File` worksheet). 

```{r create_BB16_GVA}

GVA_sectors <- GVA_sectors %>%
   dplyr::left_join(
     GVA, by = c('SIC2' = 'SIC', 'year')
   ) %>%
  dplyr::mutate(
    BB16_GVA = perc_split * GVA
  )

GVA_sectors

```

### 4. Calculate sums across sectors and years

Finally calculate the sums across all sectors and years.

```{r calc_SIC2_sums}

GVA_by_sector <- GVA_sectors %>%
  group_by(year, sector) %>%
  dplyr::summarise(
    GVA = sum(BB16_GVA)
    )

GVA_by_sector 

```

Total UK GVA is included in the `gva` dataset, this can be added to the data here:

### 5. Add in total UK GVA from `gva`

```{r extract_UK_GVA}

# Use create a table to merge in with GVA_by_sector

GVA_UK <- GVA %>%
  dplyr::filter(grepl("total.*intermediate.*", SIC)) %>%
  dplyr::transmute(
    year, 
    sector = "UK", 
    GVA
  )

GVA_UK

```

Merge this into `GVA_by_sector`:

```{r merge_GVA_by_sector_and_GVA_UK}

# Full join it together

GVA_by_sector <- GVA_by_sector %>%
  dplyr::full_join(GVA_UK)

GVA_by_sector

```

### 6. Match in Tourism data

Tourism data is provided in a separate spreadsheet and imported in the `tourism` object

```{r merge_tourism_data}

GVA_by_sector <- tourism %>%
  dplyr::transmute(year, sector = 'tourism', GVA) %>%
  dplyr::full_join(GVA_by_sector)

GVA_by_sector 
```

```{r test_merge_tourism_data}

test_that(
  'GVA_by_sector tourism values match those in tourism',
  {
    # Extract tourism gva
    
    tourism_gva <- tourism %>% 
      dplyr::arrange(year)
    
    GVA_by_sector_gva <- GVA_by_sector %>% 
      dplyr::filter(sector == 'tourism') %>% 
      dplyr::arrange(year)
        
    # Check that this merged without error.
    
    expect_equal(tourism_gva$GVA, GVA_by_sector_gva$GVA)
    
    }
  )

```

### 7. Add tourism overlap

Also need to add the `$overlap` from tourism to the `all_dcms` totals in `GVA_by_sector`

```{r remove_tourism_overlap}

GVA_by_sector <- tourism %>%
  dplyr::transmute(
    year, sector = 'all_dcms', overlap
  ) %>%
  right_join(GVA_by_sector) %>%
  dplyr::mutate(
    GVA = ifelse(!is.na(overlap), overlap + GVA, GVA)
    ) %>%
  dplyr::select(-overlap)


GVA_by_sector

```

### 8. Build the `GVA_by_sector` dataframe expected by the eesectors package

```{r build_GVA_dataframe}

# Build df to match eesectors::GVA_by_sector_2016

GVA_by_sector <- GVA_by_sector %>%
  # The 2016 SFR only inc
  dplyr::filter(
    year %in% 2010:2015
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    # Round values
    GVA = round(GVA, 2),
    # Convert sector to factor
    sector = factor(sector),
    # Convert year to integer
    year = as.integer(year)
  ) %>%
  dplyr::select(
    sector, year, GVA
  ) %>%
  dplyr::arrange(year, sector)

GVA_by_sector

```

### 9. Test the dataframe against the `eesectors::GVA_by_sector_2015`

```{r create_2016_test_cases}

# Tourism comes from Tourism tab in excel workbook

# Abs data not sue for some SIC codes (SIC 91)

test_case_2016 <- GVA_by_sector_2016 %>% 
  dplyr::mutate(GVA = round(GVA)) %>%
  dplyr::arrange(sector)
  
```

```{r test_against_2016}

test_that(
  'calculated values match those in the SFR.',
  {
      # Some general checks about the object
    
    expect_is(GVA_by_sector, c("tbl_df", "tbl", "data.frame"))
    expect_equal(nrow(GVA_by_sector), 54)
    expect_equal(levels(eesectors::GVA_by_sector_2016$sector), levels(GVA_by_sector$sector))
    
    test_case_new <- GVA_by_sector %>%
      dplyr::arrange(sector, year)
    
    test_case_2016 <- eesectors::GVA_by_sector_2016 %>%
      dplyr::arrange(sector, year)
    
    # Check equality of GVA within a given tolerance
    
    expect_equal(test_case_new$GVA, test_case_2016$GVA, tolerance = 1)
    
    # Check that years are equal
    
    expect_equal(test_case_new$year, test_case_2016$year)
    
    # Check that sectors are equal
    
    expect_equal(test_case_new$sector, test_case_2016$sector)
    }

  )

# If all good, save the dataframe out

GVA_by_sector %>% saveRDS('GVA_by_sector.Rds')
```

### 10. Print differences between `GVA_by_sector` and `eesectors::GVA_by_sector_2015`

Create a table of comparisons to see all values which do not match perfectly (due to rounding errors).

```{r print_nonmatching_values}

test <- dplyr::full_join(GVA_by_sector, (eesectors::GVA_by_sector_2016 %>% dplyr::rename(GVA_2016 = GVA)))

test %>%
  dplyr::mutate(abs_diff = abs(GVA - GVA_2016)) %>%
  dplyr::filter(abs_diff > 0) %>%
  knitr::kable()
```
